---
title: "assignment"
author: "Juli Furjes"
output: html_document
date: "2023-05-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

price data: https://www.gov.uk/government/statistical-data-sets/price-paid-data-downloads
postcode data: https://www.doogal.co.uk/Counties?county=E11000019

```{r}
library(tidyverse)
library(sf)
library(spatstat)
library(ggplot2)
library(sp)
library(leaflet)
library(geosphere)
library(lubridate)
library(hms)
library(spdep)
library(mapview)
```

## Loading data

### Cleaning data

```{r}
# CRIME DATA

# reading csv
crime_data <- read_csv("data/2023-03-west-yorkshire-street.csv")
head(crime_data)

# removing nas
crime_data <- crime_data[!is.na(crime_data$Longitude),]
crime_data <- crime_data[!is.na(crime_data$Latitude),]

# HOUSING DATA

# reading csv (without column names)
housing_data <- read.csv("data/pp-monthly-update-new-version.csv", header=FALSE)

# keeping only the columns we need
housing_data <- housing_data[ , c("V2", "V3", "V4", "V14")]

# renaming columns
colnames(housing_data) <- c("price", "date", "postcode", "county")
head(housing_data)

# print all the counties
counties <- unique(housing_data$county)
sort(counties)

# filter for West Yorkshire
housing_data <- subset(housing_data, county == 'WEST YORKSHIRE')
housing_data

# add coordinates, based on the postcode
postcodes <- read_csv("data/West Yorkshire postcodes.csv")
postcodes <- postcodes[ , c("Postcode", "Latitude", "Longitude")]
head(postcodes)

# renaming column name, so it would match with the other one
names(postcodes)[names(postcodes) == "Postcode"] <- "postcode"

# merge dataframes
housing_data <- merge(housing_data, postcodes, by="postcode")
```

### Visualizing the rates of the different types of crimes

```{r}
crime_rates <- crime_data %>% 
  group_by(`Crime type`)  %>% 
  summarise(total_count=n(),
            .groups = 'drop')

crime_rates

crime_plot <- ggplot(data=crime_rates, aes(x=`Crime type`, y=total_count)) +
  geom_bar(stat="identity", fill="slateblue", alpha=0.5) +
    xlab("Crime type") +
    ylab("Total count") +
    theme_light() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
crime_plot
```

### Filtering data

```{r}
# subsetting crime with drugs
crime_data <- crime_data[crime_data$`Crime type` == "Drugs",]

# subtracting only the year from the date
housing_data$date <- substr(housing_data$date, 1, 4)

# subtracting only the ones that were sold in 2023
housing_data <- housing_data[housing_data$date > 2022,]

# finding an area with not too much data to visualize
housing_data$postal_area <- substr(housing_data$postcode, 1, 2)

postal_rates <- housing_data %>% 
  group_by(postal_area)  %>% 
  summarise(total_count=n(),
            .groups = 'drop')

postal_rates

crime_data$area <- gsub( " .*$", "", crime_data$`LSOA name`)

crime_area_rates <- crime_data %>% 
  group_by(area)  %>% 
  summarise(total_count=n(),
            .groups = 'drop')

crime_area_rates

# filtering for HX / Calderdale
crime_data <- crime_data[crime_data$area == "Calderdale",]
housing_data <- housing_data[housing_data$postal_area == "HX",]
```


### Converting to sf

```{r}
housing_spatial <- st_as_sf(housing_data, coords = c("Longitude", "Latitude"), 
                 crs = 4326, agr = "constant")

crime_spatial <- st_as_sf(crime_data, coords = c("Longitude", "Latitude"), 
                 crs = 4326, agr = "constant")
```

### Create a dataframe from the coord of crimes and their center

```{r}
columns = c("lon","lat") 
crime_coord = data.frame(crime_data$Longitude, crime_data$Latitude) 
colnames(crime_coord) = columns

crime_coord$lon <- as.numeric(as.character(crime_coord$lon))
crime_coord$lat <- as.numeric(as.character(crime_coord$lat))

# define the center of Calderdale
center <- c(-1.858460, 53.716157)
```

## Mapping data

### Visualising crime locations

```{r}
# use this if you want the datapoints to be clustered
# leaflet() %>%
#   addTiles() %>%
#   addProviderTiles("Esri.WorldImagery", 
#                    options = providerTileOptions(opacity=0.5)) %>% 
#   setView(lng = center_point[1], lat = center_point[2], zoom = 12) %>% 
#   addCircleMarkers(
#     lat = crime_coord$lat,
#     lng = crime_coord$lon,
#     clusterOptions = markerClusterOptions()
#   )

leaflet() %>%
  addTiles() %>%
  addProviderTiles("Esri.WorldImagery", 
                   options = providerTileOptions(opacity=0.5)) %>% 
  setView(lng = center_point[1], lat = center_point[2], zoom = 12) %>% 
  addCircleMarkers(
    lat = crime_coord$lat,
    lng = crime_coord$lon
  )
```

### Visualising crime locations compared to housing prices

```{r}
getColor <- function(housing_data) {
  sapply(housing_data$price, function(price) {
    if(price < quantile(housing_data$price, probs = 0.3)) {
      "green"
    } else if(price > quantile(housing_data$price, probs = 0.3) & price < quantile(housing_data$price, probs = 0.6)) {
      "orange"
    } else {
      "red"
    } })
}

price_icons <- awesomeIcons(
  icon = 'ios-close',
  iconColor = 'black',
  library = 'ion',
  markerColor = getColor(housing_data)
)

crime_icon <- makeIcon(
  iconUrl = "https://img.freepik.com/free-icon/placeholder_318-220133.jpg?w=2000",
  iconWidth = 38, iconHeight = 40,
  iconAnchorX = 22, iconAnchorY = 94
)

# use this if you want the datapoints to be clustered
# leaflet() %>%
#   addTiles() %>%
#   addProviderTiles("Esri.WorldImagery", 
#                    options = providerTileOptions(opacity=0.5)) %>% 
#   setView(lng = center_point[1], lat = center_point[2], zoom = 12) %>% 
#   addMarkers(
#     lat = crime_coord$lat,
#     lng = crime_coord$lon,
#     icon = crime_icon,
#     clusterOptions = markerClusterOptions()
#   ) %>%
#   addAwesomeMarkers(housing_data$Longitude, housing_data$Latitude, icon = price_icons,
#     clusterOptions = markerClusterOptions())

leaflet() %>%
  addTiles() %>%
  addProviderTiles("Esri.WorldImagery", 
                   options = providerTileOptions(opacity=0.5)) %>% 
  setView(lng = center_point[1], lat = center_point[2], zoom = 12) %>% 
  addMarkers(
    lat = crime_coord$lat,
    lng = crime_coord$lon,
    icon = crime_icon
  ) %>%
  addAwesomeMarkers(housing_data$Longitude, housing_data$Latitude, icon = price_icons)
```
### Visualizing data occurences from both of the datasets

```{r}
# filtering for the cheap housings (bottom 20%)
cheap_housing_data <- housing_data[housing_data$price < quantile(housing_data$price, probs = 0.3),]

cheap_housing_spatial <- st_as_sf(cheap_housing_data, coords = c("Longitude", "Latitude"), crs = 4326, agr = "constant")

# visualizing
map_housing <-  mapview(cheap_housing_spatial, map.types = "OpenStreetMap", 
        col.regions = "#32612D",
        color = "white",
        legend = TRUE, 
        layer.name = "Sold cheap houses in 2023",
        homebutton = FALSE, lwd = 0.5) 

map_crime <-  mapview(crime_spatial, 
          map.types = "OpenStreetMap", 
        col.regions = "#32612D", 
        color = "white",
        legend = TRUE, 
        layer.name = "Crime locations from 2023-03",
        homebutton = FALSE, lwd = 0.5)

leafsync::sync(map_housing, map_crime)
```

## Calculating nearest neighbour

```{r}
# creating three pricing groups
for (i in 1:length(housing_data$price)) {
  if(housing_data$price[i] < quantile(housing_data$price, probs = 0.3)) {
      housing_data$price_category[i] <- "low"
    } else if(housing_data$price[i] > quantile(housing_data$price, probs = 0.3) & housing_data$price[i] < quantile(housing_data$price, probs = 0.6)) {
      housing_data$price_category[i] <- "medium"
    } else {
      housing_data$price_category[i] <- "high"
    }
}

housing_low <- housing_data[housing_data$price_category == "low",]
housing_med <- housing_data[housing_data$price_category == "medium",]
housing_high <- housing_data[housing_data$price_category == "high",]

#define the radius
radius <- 10000  # Radius length

# create the window
window <- disc(radius = radius, centre = center)

# visualize window
leaflet() %>%
  addTiles() %>%
  addProviderTiles("Esri.WorldImagery", 
                   options = providerTileOptions(opacity=0.5)) %>% 
  setView(lng = center_point[1], lat = center_point[2], zoom = 12) %>% 
  addMarkers(
    lat = crime_coord$lat,
    lng = crime_coord$lon,
    icon = crime_icon
  ) %>%
  addAwesomeMarkers(housing_data$Longitude, housing_data$Latitude, icon = price_icons) %>%
  addCircles(lng = center[1], lat = center[2], radius = radius) 

# create ppp objects based on the window and house price categories
crime_ppp <- ppp(x = crime_data$Longitude, y = crime_data$Latitude, window = window)
lowprice_ppp <- ppp(x = housing_low$Longitude, y = housing_low$Latitude, window = window)
medprice_ppp <- ppp(x = housing_med$Longitude, y = housing_med$Latitude, window = window)
highprice_ppp <- ppp(x = housing_high$Longitude, y = housing_high$Latitude, window = window)

# calculate neaarest neighbour
nndistances_low <- nndist(lowprice_ppp, crime_ppp)
nndistances_med <- nndist(medprice_ppp, crime_ppp)
nndistances_high <- nndist(highprice_ppp, crime_ppp)

# plot histograms
hist(nndistances_low)
hist(nndistances_med)
hist(nndistances_high)

# print mean distance
print(paste('Mean distance to a drug crime scene with low house prices: ', mean(nndistances_low)))

print(paste('Mean distance to a drug crime scene with medium house prices: ', mean(nndistances_med)))

print(paste('Mean distance to a drug crime scene with high house prices: ',mean(nndistances_high)))
```

### Checking whether the spatial pattern of crimes is random

```{r}
qtest <- quadrat.test(crime_ppp)
qtest

plot(qtest)
```

### Testing the significance of the neighbour results

```{r}
# creating dataframe for the testing
nndistances_low_df <- data.frame(
  category = "low",
  mean_dist = nndistances_low
)

nndistances_med_df <- data.frame(
  category = "medium",
  mean_dist = nndistances_med
)

nndistances_high_df <- data.frame(
  category = "high",
  mean_dist = nndistances_high
)

nndistances_df <- rbind(nndistances_low_df, nndistances_med_df, nndistances_high_df)

# boxplot
nndistances_df %>% 
  mutate(category = factor(category, levels=c("low", "medium", "high"))) %>%
  ggplot(aes(x=category, y=mean_dist)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + 
    xlab("Price Category") +
    ylab("Mean Distance from Crime Scene") +
    theme_light()

# one-way anova
one.way <- aov(mean_dist ~ category, data = nndistances_df)
summary(one.way)
```
